DRAFT: Factors Influencing Abundance & Virus Isolation in Mosquitoes

### 2024-03-07 R/processing-code/MADAproject_pt2_ruiz.qmd
Fixed manuscript render error
loaded package and inserted image generated in processing qmd

```{r trap-data-import}
# Construct the path to the Excel file
file_path_mos <- here("data", "processed-data", "MOSQ_MADA.xlsx")

# import the data to R and store it in a dataframe named "mosquito"
mosquito_raw <- read_excel(file_path_mos)

>>>>

```{r trap-data-import}
# Construct the path to the Excel file
#file_path_mos <- here("data", "raw-data", "MOSQ_MADA.xlsx")
file_path_mos <- here("data", "raw-data", "MOSQ_MADA.xlsx")
file_exists_check <- file.exists(file_path_mos)
file_path_mos <- here("data", "raw-data", "MOSQ_MADA.xlsx")

file_exists_check <- file.exists("/Users/andrewruiz/MADA_course/RUIZ-MADA-project/data/raw-data/MOSQ_MADA.xlsx")
print(file_exists_check)  # Should return TRUE if the path is correct
print(file_path_mos)

print(file_exists_check)  # This should return TRUE if the file path is correct
# import the data to R and store it in a dataframe named "mosquito"
mosquito_raw <- read_excel(file_path_mos)

### 2024-03-07 products/manuscript/Manuscript.qmd
Fixed manuscript render error
loaded package and inserted image generated in processing qmd

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(readxl)

@@

| Data                         | Description                                                                                                                                           | Source                                |
|------------------|-------------------------------|-----------------------|
| Mosquito Surveillance Data   | Records of mosquito species counts, locations, and dates of collection. Data on the presence of EEE in mosquito populations is particularly valuable. | XXX County Mosquito Control           |
| Weather Data                 | Temperature, humidity, rainfall, and wind speed influence mosquito activity, population dynamics, and virus transmission.                             | NOAA                                  |
| Bird Population Data         | Since birds are a natural reservoir for EEE, information on bird populations and migration patterns could be relevant.                                | eBird                                 |
| Human Case Reports           | Data on confirmed human cases of EEE, including location, date of onset, and clinical outcomes.                                                       | MA Dept of Public Health publications |
| Veterinary Surveillance Data | Since horses and other mammals can also be affected by EEE, veterinary records could provide early warning signs of virus activity in an area.        | MA Dept of Public Health              |

@@

One factor to consider is if early virus isolations are associated with
cases in humans and other mammals. The table below shows the date of the
first EEE virus isolation in mosquitoes and the number of human of other
animal cases by year.

![](images/eee_isolation_cases-01.png){fig-align="center" width="320"}

@@

## Basic statistical analysis
*These will be generated in later phases of the exercise*

@@

*These will be generated in later phases of the exercise* 
# Discussion *this section will be completed when analysis is done*

## Strengths and Limitations 
*this section will be updated after analysis is complete*

@@

## Conclusions *this section will be completed when analysis is done*
{{< pagebreak >}}

>>>>

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(readxl)
library(dplyr)
library(kableExtra)
library(dplyr)
library(readr)

@@

| Data                         | Description                                                                                                                                           | Source                                |
|------------------------------|-------------------------------------------------------------------------------------------------------------------------------------------------------|---------------------------------------|
| Mosquito Surveillance Data   | Records of mosquito species counts, locations, and dates of collection. Data on the presence of EEE in mosquito populations is particularly valuable. | XXX County Mosquito Control           |
| Weather Data                 | Temperature, humidity, rainfall, and wind speed influence mosquito activity, population dynamics, and virus transmission.                             | NOAA                                  |
| Bird Population Data         | Since birds are a natural reservoir for EEE, information on bird populations and migration patterns could be relevant.                                | eBird                                 |
| Human Case Reports           | Data on confirmed human cases of EEE, including location, date of onset, and clinical outcomes.                                                       | MA Dept of Public Health publications |
| Veterinary Surveillance Data | Since horses and other mammals can also be affected by EEE, veterinary records could provide early warning signs of virus activity in an area.        | MA Dept of Public Health              |

@@

It is important to know where the mosquitoes infected with EEE are
captured. Below is a table showing the virus isolations for EEE in
trapped mosquitoes by county between 2014 and 2020.

![](images/eee_isolation_county.png){fig-align="center"}

The geographic focus for this study is Bristol County. One factor to
consider is if early virus isolations are associated with cases in
humans and other mammals. The table below shows the date of the first
EEE virus isolation in mosquitoes and the number of human of other
animal cases by year.

![](images/eee_isolation_cases-03.png){fig-align="center"}

@@

## Basic statistical analysis

*These will be generated in later phases of the exercise*

@@

*These will be generated in later phases of the exercise* \# Discussion
*this section will be completed when analysis is done*

## Strengths and Limitations

*this section will be updated after analysis is complete*

@@

## Conclusions *this section will be completed when analysis is done*
